---
title: "Homework 8 - STAT 5361 Statistical Computing"
author:
  - Sen Yang^[<sen.2.yang@uconn.edu>; M.S. student at
    Department of Statistics, University of Connecticut.]
date: "`r format(Sys.time(), '%d %B %Y')`"
documentclass: article
papersize: letter
fontsize: 11pt
bibliography: template.bib
biblio-style: asa
output:
  bookdown::pdf_document2
abstract: |
    This is homework 8 for STAT 5361 - Statistical Computing.
---


```{r setup, echo = FALSE, message = FALSE, warning = FALSE}
## some utility functions, see the source code for details
source("utils_template.R")

## specify the packages needed
pkgs <- c("DT", "leaflet", "splines2", "webshot")
need.packages(pkgs)

## external data can be read in by regular functions,
## such as read.table or load

## for latex and html output
isHtml <- knitr::is_html_output()
isLatex <- knitr::is_latex_output()
latex <- ifelse(isLatex, '\\LaTeX\\', 'LaTeX')

## specify global chunk options
knitr::opts_chunk$set(fig.width = 5, fig.height = 4, dpi = 300,
                      out.width = "90%", fig.align = "center")

```


# Orstein–Uhlenbeck Process {#sec:5.3.3}

## Transition Distribution

SDE of Ornstein-Uhlenbeck process is
$$\mathrm{d}r(t)=\alpha(b-r(t))\mathrm{d}t+\sigma\mathrm{d}W(t)$$
Let $H(t)=\int_0^{t}-\alpha\mathrm{d}s=-\alpha t$. 
By Ito's formula,
$$\mathrm{d}(e^{\alpha t}r(t))=e^{\alpha t}\alpha b\mathrm{d}t+e^{\alpha t}\sigma\mathrm{d}W(t)$$
That is,
\begin {align*}
    \int_0^{t}\mathrm{d}(e^{\alpha t}r(t)) &= \int_0^{t} e^{\alpha s}\alpha 
    b\mathrm{d}s+\int_0^{t} e^{\alpha s}\sigma\mathrm{d}W(s) \\
    e^{\alpha t}r(t)-e^{0}r(0) &= \alpha b\int_0^{t} e^{\alpha s}\mathrm{d}s+\sigma
    \int_0^{t} e^{\alpha s}\mathrm{d}W(s) \\
    r(t) &= e^{-\alpha t}r(0)+ b(1-e^{-\alpha t})+e^{-\alpha t}
    \sigma\int_0^{t} e^{\alpha s}\mathrm{d}W(s) \\
    e^{-\alpha\Delta}r(t) &= e^{-\alpha (t+\Delta)}r(0)+ b(e^{-\alpha\Delta}-e^{-\alpha 
    (t+\Delta)})+e^{-\alpha (t+\Delta)} \sigma\int_0^{t} e^{\alpha s}\mathrm{d}W(s) 
\end {align*}

Since $$r(t+\Delta) = e^{-\alpha (t+\Delta)}r(0)+ b(1-e^{-\alpha (t+\Delta)})+e^{-\alpha 
    (t+\Delta)} \sigma\int_0^{(t+\Delta)} e^{\alpha s}\mathrm{d}W(s),$$
Then,
\begin {align*}
    r(t+\Delta) &= e^{-\alpha\Delta}r(t)+b(1-e^{-\alpha\Delta})+e^{-\alpha 
    (t+\Delta)}\sigma\int_t^{(t+\Delta)} e^{\alpha s}\mathrm{d}W(s) \\
    &= e^{-\alpha\Delta}r(t)+b(1-e^{-\alpha\Delta})+e^{-\alpha 
    (t+\Delta)}\sigma\sqrt{\int_t^{(t+\Delta)} e^{2\alpha s}\mathrm{d}s}Z \\
    &= e^{-\alpha\Delta}r(t)+b(1-e^{-\alpha\Delta})+e^{-\alpha
    (t+\Delta)}\sigma\sqrt{\frac{e^{2\alpha(t+\Delta)}-e^{2\alpha}}{2\alpha}}Z \\
    &= e^{-\alpha\Delta}r(t)+b(1-e^{-2\alpha\Delta})+\sigma\sqrt{\frac{1-e^
    {-\alpha\Delta}}{2\alpha}}Z \\
\end {align*}

## A Random Walk for the Process

The algorithm to implement a random walk construction for the process:

\begin{algorithm} [H]
\caption{Implement a random walk construction for the process}
\begin{algorithmic} 
\STATE Step 1: Set $r(0)=1, i=1,\Delta=1/500$.
\STATE Step 2: Sample $Z_i\sim N(0,1) $.
\STATE Step 3: compute $r(i)$.
\STATE Step 4: $i=i+\Delta$.
 \IF {$i>\frac{500-1}{1/500}=249500$}
 \STATE Break the loop.
 \ELSE
 \STATE Return Step 2.
 \ENDIF
\end{algorithmic}
\end{algorithm}

Realize this algorithm in R.
```{r walk, echo=T, eval=F}
# Orstein–Uhlenbeck Process
alpha <- c(0.1,1,5)
sigma <- c(0.1,0.2,0.5)
b <- c(-5,5)

## random walk
ran_walk <- function(alpha,sigma,b,r0=1,T=500,delta=1/500) {
  dim<-length(alpha)*length(sigma)*length(b)
  results<-data.frame(matrix(0,(T-r0)/delta,dim*5))
  for (i in 1:dim) {results[,i*5-1]<-rnorm(nrow(results))}
  for (j in 1:length(alpha)) {
    for (k in 1:length(sigma)) {
      for (l in 1:length(b)) {
        num <- (j-1)*6+(k-1)*2+l
        results[,num*5-4]<-alpha[j]
        results[,num*5-3]<-sigma[k]
        results[,num*5-2]<-b[l]
      }
    }
  }
  ri<-data.frame(matrix(r0,1,dim))
  for (i in 1:nrow(results)) {
    ri<-exp(-results[i,(1:dim)*5-4]*delta)*ri[1,1:dim]+
      results[i,(1:dim)*5-2]*(1-exp(-2*results[i,(1:dim)*5-4]*delta))+
      results[i,(1:dim)*5-3]*sqrt((1-exp(-results[i,(1:dim)*5-4]*delta))/2/
             results[i,(1:dim)*5-4])*results[i,(1:dim)*5-1]
    results[i,(1:dim)*5]<-ri[1:dim]
  }
  results
}
results<-ran_walk(alpha,sigma,b)

## plot
res2<-results[,(1:dim)*5]
res2<-cbind(col(1:nrow(results)),res2)
library(ggplot)
for (j in 1:length(alpha)) {
  for (k in 1:length(sigma)) {
    for (l in 1:length(b)) {
      num <- (j-1)*6+(k-1)*2+l
      ggplot() + deom_line(data=res2,aes(x=res2[,1],y=res2[,num])) + 
        labs(x="t",y="r(t)",
             title=expression(paste(alpha,"=",alpha[j],", ",sigma,"=",sigma[k],", b=",b[l]))) + 
        theme(plot.title = element_text(hjust = 0.5))
    }
  }
}
```

## Simulation of the Process by Euler–Maruyama method
# Poisson Process
## Distribution of $N(5)$

Given $T>0$, $Z=\int_0^T\lambda(t)\mathrm{d}t=\int_0^T\sqrt{t}+e^{-t}\sin(2\pi t)\mathrm{d}t$.
Calculate this integral by R,

``` {r inhomo_para, echo=T}
## Integration
lamda_t <- function(t) {
  lamdat<- sqrt(t) + exp(-t)*sin(2*pi*t)
  lamdat
}

Z <- integrate(lamda_t,0,5)
Z$value
```

Therefore, $N(5)\sim \mathrm{Poisson}(Z), where\;\;Z=7.607738.$

## Function to Simulate from this Poisson Process in R

Since $\lambda(t)=\sqrt{t}+e^{-t}\sin(2\pi t)\le\sqrt{t}+e^{-t}$, we choose $\lambda_0(t)=\sqrt{t}+e^{-t}$.
Therefore,
$$\Lambda(\tau)=\int_0^{\tau}\sqrt{t}+e^{-t}\mathrm{d}t=\frac{2}{3}\tau^{2/3}-e^{-tau}+1$$

The algorithm to simulation from this Poisson process:
\begin{algorithm} [H]
\caption{Simulation from a Poisson process}
\begin{algorithmic} 
\STATE Step 1: Derive mean function $\Lambda(\tau)$ by $\Lambda(\tau)=\int_0^{\tau}\lambda(t)\mathrm{d}t$
\STATE Step 2: Generate $S_1,S_2,...,$ from a homogeneous Poisson process with rate one
\STATE Step 3: Let $T_i=\Lambda^{-1}(S_i), i=1,2,...$
\STATE Step 4: Sample $U \sim Unif(0,1)$, for each $i$
\IF {$U < \lambda(T_i)/\lambda_0(T_i)$}
\STATE Return $T_i$
\ELSE 
\STATE Go to Step 3.
\ENDIF
\end{algorithmic}
\end{algorithm}

Build corresponding funcion in R.

``` {r poi_prc, echo=T}
## poisson process
tau_t<-matrix(0,1,1)
for (i in 1:1000) {
  S<-5* runif(rpois(1,5))
  T<-matrix(0,length(S),1)
for (j in 1: length(S)) {
  if (length(S)==0) next
  lamda_f <- function(t) {
    lamda_f <- 2/3*t^(3/2)-exp(-t)+1-S[j]
    lamda_f
  }
  T[j,1]<-uniroot(lamda_f,c(0,5))$root
  tau<-matrix(0,length(S),1)
}
  for (k in 1:nrow(T)) {
    if (length(T)==0) next
    u0 <- runif(1)
    if (u0<=(sqrt(T[k,1])+exp(-T[k,1])*sin(2*pi*T[k,1]))/(sqrt(T[k,1])+exp(-T[k,1]))) {
      tau[k,1]<-T[k,1]
    }
    else next
  }
  tau_t<-rbind(tau_t,tau)
}
tau_t<-as.data.frame(tau_t[tau_t>0 & tau_t<5])
```

## Poisson Process Simulation
Generate events from this Poisson process 1000 times and plot the results.

``` {r poi_sim, echo=T}
true_f <- function(t) {
  true_f<- (sqrt(t)+exp(-t)*sin(2*pi*t))/(Z$value)
  true_f
}
library(ggplot2)
ggplot() + stat_function(aes(col="true"), fun=true_f) +
  geom_density(data=tau_t,aes(tau_t[,1], col='kernel')) + 
  scale_color_manual(values = c("red", "blue")) +
  labs(title = paste("kernel density vs. true density with "),
       x = "t", y = "value") + theme(plot.title = element_text(hjust = 0.5))
```





